---
title: "Demographics"
output: 
    html_document:
      code_download: TRUE
      toc: TRUE
      toc_float:
        collapsed: FALSE
      toc_depth: 1
      code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r data prep, echo = FALSE, warning = FALSE, message = FALSE, error = FALSE}
# Loading packages
library(psych)
# library(sjPlot)
library(effects)
library(magrittr)
library(parameters)
library(dplyr)
library(tidyr)
library(rio)
library(ggplot2)
library(corrplot)
library(factoextra)

# Function to clean document
source("functions/cleaning.R")
source("functions/demographics counts.R")

# Setting global chunk options
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)

options(scipen = 999)

# Importing data
wide_raw <- import("data/pilot_raw_removed4.csv") 

# Cleaning data using function
clean_data <- get_clean_data(wide_raw)
```

# Demographics {.tabset .tabset-fade .tabset-pills}

## Age

```{r}
clean_data %>%
  select(subid, age) %>% 
  na.omit() %>% 
  summarize(mean = mean(age),
            sd = sd(age))
```

## Gender

```{r}
gender_counts <- clean_data %>% 
  select(subid, gender) %>%  
  na.omit() %>% 
  unique() %>% 
  group_by(gender) %>% 
  count()

dem_percent_table(gender_counts)
```

Everyone responded to this question (211 when you add it up)

## Race/Ethnicity

```{r}
ethnicity_counts <- clean_data %>% 
  select(subid, race) %>%  
  na.omit() %>% 
  unique() %>% 
  group_by(race) %>% 
  count()

dem_percent_table(ethnicity_counts)
```

Everyone also responded to this question

## Education - Participant

```{r}
edu_counts <- clean_data %>% 
  select(subid, education) %>%  
  na.omit() %>% 
  unique() %>% 
  group_by(education) %>% 
  count()

dem_percent_table(edu_counts)
```

## Education - Parents

```{r}
edu_parent_counts <- clean_data %>% 
  select(subid, parent_education) %>%  
  na.omit() %>% 
  unique() %>% 
  group_by(parent_education) %>% 
  count()

dem_percent_table(edu_parent_counts)
```

## Birth country

```{r}
birth_country_counts <- clean_data %>% 
  select(subid, country_birth) %>% 
  na.omit() %>% 
  unique() %>% 
  group_by(country_birth) %>% 
  count()

dem_percent_table(birth_country_counts)
```

## Country raised

```{r}
raised_country_counts <- clean_data %>% 
  select(subid, country_raised) %>% 
  na.omit() %>% 
  unique() %>% 
  group_by(country_raised) %>% 
  count()

dem_percent_table(raised_country_counts)
```

### Follow-up years in US

```{r}
clean_data %>% 
  filter(country_raised != "Raised in US") %>% 
  select(raised_follow)
```

Should we removed the 5 primarily raised in another country? They all are US citizens and have lived in the US for 12-35 years. Seems like they have a good handle on US politics.

## Native language

```{r}
lang_counts <- language_counts <- clean_data %>% 
  select(subid, native_language) %>% 
  na.omit() %>% 
  unique() %>% 
  group_by(native_language) %>% 
  count()

dem_percent_table(lang_counts)
```

### Native language follow-up

```{r}
clean_data %>% 
  filter(native_language == "Not English") %>% 
  select(native_lang_follow)
```

Everyone has spoken for almost 20 years or more, so doesn't seem like we need to filter anyone out.