---
title: "Pilot Analyses"
output: 
    html_document:
      code_download: TRUE
      toc: TRUE
      toc_float:
        collapsed: FALSE
      toc_depth: 1
      code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r data prep, echo = FALSE, warning = FALSE, message = FALSE, error = FALSE}
# Loading packages
library(psych)
library(effects)
library(magrittr)
library(parameters)
library(dplyr)
library(tidyr)
library(rio)
library(ggplot2)
library(corrplot)
library(factoextra)
library(emmeans)

# Function to clean document
source("functions/cleaning.R")

# Setting global chunk options
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)

options(scipen = 999)

# Importing data
wide_raw <- import("data/pilot_raw_removed4.csv") 

# Cleaning data using function
clean_data <- get_clean_data(wide_raw)
```

# Descriptives {.tabset .tabset-fade .tabset-pills}

## Overall

```{r}
descrip_overall <- psych::describe(clean_data)
descrip_overall
```

## Sample size

```{r}
clean_data %>% 
  select(subid) %>% 
  unique() %>% 
  nrow()
```

### Per condition

```{r}
clean_data %>% 
  select(subid, condition) %>% 
  unique() %>% 
  group_by(condition) %>% 
  count()
```

A lot were removed from the competitive target in comparison to the other conditions.

## Histograms

### Distance in coffee shop

```{r}
hist(clean_data$distance_coffee)
````

Seat E is option "5," or the farthest seat away. It is left/negative-skewed to this option.

#### Is the skew non-normal?

```{r}
se_skew <- sqrt((6*211*(211-1))/((211-2)*(211+1)*(211+3)))
z_coffee <- descrip_overall$skew[13]/se_skew
z_coffee
```

According to Kim (2013; https://www.ncbi.nlm.nih.gov/pmc/articles/PMC3591587/), a sample of 50 < n < 300 with an absolute z-score of >3.29 is non-normal. This z-score of -2.42 is less than 3.29, making the skew normal. 

Formula used for se of skew:
se_skew = sqrt(6*N*(N-1) / ((N-2)*(N+1)*(N+3)))

https://www.ibm.com/support/pages/standard-errors-skewness-and-kurtosis-are-all-same-set-variables

Committee member (Sara W) also informed me that unless I was missing responses for one answer entirely on one side, the skew is not something to be concerned about. I am keeping this test here as a reference for future, but am removing other z tests for skew below because none meet the observational skew test Sara suggested.

### Distance in town

```{r}
hist(clean_data$distance_town)
````

Right skewed (not as far of a distance), kind of plateaus at option 5. Still normal skew.

### Realistic threat

```{r}
hist(clean_data$realistic_q)
````

Left/negative-skewed, with 4 = "Extremely threatening." I don't think this is concerning because it shows people are experiencing more threat overall and we had more threatening conditions than not. Skew is normal.

### Symbolic threat

```{r}
hist(clean_data$symbolic_q)
````

Left/negative-skewed, same as with realistic threat above. Skew is normal.

### Explicit threat posed by target

```{r}
hist(clean_data$explicit_targ)
````

Looks evenly spread out.

### Explicit threat posed by political out-group 

```{r}
hist(clean_data$explicit_group)
````

Left/negative skew. Seems people are more willing to discuss explicit threat of the group than the individual. Skew is normal.

### Political orientation

#### Overall

```{r}
hist(clean_data$pol_orient_1)
````

Right/positive-skewed, not surprising as filtered for liberals and 1 = "extremely liberal." Skew is normal.

#### Social 

```{r}
hist(clean_data$pol_orient_2)
````

Right/positive-skewed, still not surprised for the same reasons. Skew is normal.

#### Economic

```{r}
hist(clean_data$pol_orient_3)
````

Right/positive-skewed. Skew is normal.

# Scale validation {.tabset .tabset-fade .tabset-pills}

## Correlation matrix

```{r}
# Getting only the variables in order to do an EFA
efa_data <- clean_data %>% 
  select(distance_coffee, distance_town, realistic_q, symbolic_q)

cor_matrix <- cor(efa_data)
cor_matrix
```

## EFA

### Can an EFA be run? 

```{r EFA check}
# Check factor structure to see if there is enough correlation, etc. for factor analysis
check_factorstructure(efa_data)
```

### What number of factors are advisable?

```{r EFA n factor check}
# What is the most recommended number of EFA factors?
factors_exp <- n_factors(efa_data)
factors_exp

# How many factors are supported by what number of methods?
summary(factors_exp)

nfactors(efa_data)
```

It appears that 1 factor (as we planned) or maybe 2 factors are most advisable!

### 1 Factor EFA

```{r EFA 1 fac}
efa_1fac <- psych::fa(efa_data, nfactors = 1)

efa_1fac %>% 
  model_parameters(sort = TRUE, threshold = "max")
```

With the correlations not being as high for the distance measures and their respective factor loadings also being low on the one factor EFA, it appears that 1 factor with just realistic and symbolic threat is most advisable. The distance measures will be explored separately.

## Cronbach's alpha

### Alpha with realistic and symbolic

```{r}
efa_itt <- efa_data %>% 
  select(-distance_town, -distance_coffee)

psych::alpha(efa_itt)
```

Intergroup threat theory (realistic and symbolic threat) measures are highly reliable, which is consistent with correlations and EFA results.

```{r}
efa_distance <- efa_data %>% 
  select(distance_town,  distance_coffee)

psych::alpha(efa_distance)
```

Distance measures not reliable together, which seems consistent with EFA results. They represent two separate factors. They will each need to be tested individually.

## Current Thoughts

- Primary measure of threat will be a composite (average) of realistic and symbolic threat
- All analyses will also be predicted by distance measures to see how they vary; same goes for explicit threat measures

# Relationship with explicit threat {.tabset .tabset-fade .tabset-pills}

```{r}
threat_all <- clean_data %>% 
  select(distance_coffee, distance_town, realistic_q, symbolic_q, explicit_targ,
         explicit_group)

cor_matrix_all <- cor(threat_all)
cor_matrix_all

cor_plot_all_threat <- corrplot(cor_matrix_all, method = "number")

colnames(cor_plot_all_threat) = rep(c('Distance: Coffee', 
                                       'Distance: Town',
                                       'Realistic',
                                       'Symbolic',
                                       'Explicit: Target',
                                       'Explicit: Group'))

rownames(cor_plot_all_threat) = rep(c('Distance: Coffee', 
                                       'Distance: Town',
                                       'Realistic',
                                       'Symbolic',
                                       'Explicit: Target',
                                       'Explicit: Group'))
corrplot(cor_plot_all_threat, 
         is.corr = TRUE, 
         #method = "number", 
         method = 'color',
         col.lim = c(0, 1),
         tl.cex = .85,
         tl.col = 'black',
         addgrid.col = 'white',
         addCoef.col = 'grey50')
```

The high correlations with explicit threat, though not expected, validate our measures of symbolic and realistic threat.

# ANOVA: Conditions prediciting threat {.tabset .tabset-fade .tabset-pills}

## Implicit threat: Realistic and Symbolic

Currently using a composite supported by my EFA above, combining realistic threat, and symbolic threat. 

```{r}
clean_data %<>%   
  mutate(itt_comp = rowMeans(select(., c("realistic_q", "symbolic_q"))))
```

### Mean

```{r}
mean(clean_data$itt_comp)
```

### SD

```{r}
sd(clean_data$itt_comp)
```

### Range

```{r}
range(clean_data$itt_comp)
```

### ANOVA

#### Summary results

```{r}
contrasts(clean_data$condition)
results <- aov(itt_comp ~ condition, data = clean_data)
summary(results)
```

#### Residuals & Heteroscedastisticity

```{r}
plot(results)
```

A little weird, but working with categorical variables. The QQ especially suggests some non-normality of residuals, but regression odels are robust to violations of normality and heteroscedastisticity.

#### Post-hoc test

```{r}
emmeans(results, pairwise ~ condition)
```

Control differs from warm targ, covid targ, comp targ, and loss targ
Loss targ and warm targ also differ from one another

The warm targ and comp targ do not differ (somewhat applicable to Study 1, but not post-hoc or female vs male)

## Distance coffee

### ANOVA

#### Summary results

```{r}
results_coffee <- aov(distance_coffee ~ condition, data = clean_data)
summary(results_coffee)
```

#### Residuals

```{r}
plot(results_coffee)
```

Residuals and QQ look a little more normal than wwith the threat composite. Same reasoning here to assume regression is robst to violations.

#### Marginal means & pairwise

```{r}
emmeans(results_coffee, pairwise ~ condition)
```

Loss and covid differ from control; loss also differs from warm; warm and control did not differ

## Distance town

### ANOVA

#### Summary results

```{r}
results_town <- aov(distance_town ~ condition, data = clean_data)
summary(results_town)
```

#### Residuals

```{r}
plot(results_town)
```

QQ suggests some non-normality, but regression is robust to violations of normality.

#### Marginal means & pairwise

```{r}
emmeans(results_town, pairwise ~ condition)
```

Control and covid target and control and warm target were the only ones that significantly differed. The covid target is interesting, it may be that we anticipate different states to have different responses. Also, the fact that the warm target is the highest is also surprising. Something about this measure seems inconsistent with the others.

# Explicit measures {.tabset .tabset-fade .tabset-pills}

## EFA

```{r}
explicit_2 <- clean_data %>% 
  select(explicit_targ, explicit_group)

explicit_1fac <- psych::fa(explicit_2, nfactors = 1)

explicit_1fac %>% 
  model_parameters(sort = TRUE, threshold = "max")
```

### Alpha

```{r}
psych::alpha(explicit_2)
```

## Descriptives

### Composite explicit measure

```{r}
clean_data %<>%   
  mutate(explicit_comp = rowMeans(select(., c("explicit_targ", "explicit_group"))))
```

#### Mean

```{r}
mean(clean_data$explicit_comp)
```

#### SD

```{r}
sd(clean_data$explicit_comp)
```

#### Range

```{r}
range(clean_data$explicit_comp)
```

### Correlation with ITT composite

```{r}
cor_threat_comps <- clean_data %>% 
  select(itt_comp, explicit_comp)

cor(cor_threat_comps)
```

Explicit threat and ITT measures are highly correlated. This provides additional validation.

## ANOVA

#### Summary results

```{r}
results_explicit <- aov(explicit_comp ~ condition, data = clean_data)
summary(results_explicit)
```

#### Residuals

```{r}
plot(results_explicit)
```

QQ suggests some non-normality (as with the other measures), but regression models are often robust to violations.

#### Marginal means & pairwise

```{r}
emmeans(results_explicit, pairwise ~ condition)
```

As with the ITT measure, warm and the control differ, warm and loss differ, and loss and control differ. The control also differs from the covid target and comp target. The warm and comp target do not differ. Basically identical to the ITT results, providing extra validation.
